Last glacial cycle glacier erosion potential in the Alps

### 2021-01-28 articles/erosion/alps-erosion.tex
Rework the choice of erosion law discussion.

    The choice of erosion law significantly impacts results presented in this
    study (Fig.~\ref{fig:powerlaws}). The erosion law based on quantified
    sediment yields from Patagonian and Antarctic Peninsula tidewater glaciers
    \citet[$\dot{e} = 5.2 \times 10^{-8} u_\mathrm{b}^{2.34}$]{Koppes.etal.2015},
    our default choice, yields moderate yet strongly localized cumulative erosion
    potential (Figs~\ref{fig:cululative} and Fig.~\ref{fig:powerlaws}a).

    The erosion law based on measured suspended sediment load and composition
    in the outlet creek of Franz-Joseph Glacier
    \citep[$\dot{e} = 2.7 \times 10^{-7} u_\mathrm{b}^{2.02}$]{Herman.etal.2015},
    also yields strongly localized erosion potential but much higher values
    (Fig.~\ref{fig:powerlaws}b).

    The erosion law based on a global compilation of glacier velocity and erosion
    rates
    \citet[$\dot{e} = 1.665 \times 10^{-1} u_\mathrm{b}^{0.6459}$]{Cook.etal.2020},
    results in similarly high values, but a flatter (less localized) pattern of
    cumulative erosion potential (Fig.~\ref{fig:powerlaws}ac), with lower
    values in the troughs and more erosion in the mountains.

    These three erosion laws appear incompatible. Given the scarcity of measured
    glacier erosion rates, and the variety in glacier systems, climate and
    geologic setting, this is not a surprise.  But as more extensively
    discussed by \citet{Cook.etal.2020}, they do not have to be entirely
    incompatible. The erosion's feedback onto ice dynamics
    may yield to a delocalization of high-velocity flow in the long run.

    Alternative empirical erosion laws by
    \citet{Herman.etal.2015, Cook.etal.2020} yield glacial cycle cumulative
    erosion potential several orders of magnitude higher than modelled with the
    law by \citet{Koppes.etal.2015}, with values in the kilometres or tens of
    kilometres in places, which is of course well above the existing glacial
    relief from the Alps resulting from multiple glacial cycles.

    It is surprising that the erosion law by \citet{Koppes.etal.2015},
    calibrated on tidewater glaciers, yields more realistic erosion rates for
    the glacial Alps. However, we can imagine that Alpine paleoglaciers
    flowing on a thick sediment bed (and sometimes into proglacial lakes, not
    included), behave closer to tidewater glaciers than the Franz-Josef steep
    mountain glacier. Moreover, the cold climate of the LGM Alps resembled more
    that of modern Patagonia and Antarctica, than that of Franz-Joseph Glacier
    or most glaciers in the global compilation by \citet{Cook.etal.2020}.

@@

    \begin{figure*}
      \centerline{\includegraphics{alpero_powerlaws}}
      \caption{%
        Modelled cumulative (time-integrated) glacial erosion potential for
        three different erosion laws published by
        \textbf{(a)} \citet{Koppes.etal.2015},
          $5.2 \times 10^{-8} u_\mathrm{b} ^{2.34}$ (same as
          Fig.~\ref{fig:cumulative}a but with a different colour scale),
        \textbf{(b)} \citet{Herman.etal.2015},
          $2.7 \times 10^{-7} u_\mathrm{b} ^{2.02}$, and
        \textbf{(c)} \citet{Cook.etal.2020},
          $1.665 \times 10^{-1} u_\mathrm{b} ^{0.6459}$.
        \textbf{(d)} Corresponding erosion power laws, and
        \textbf{(e)} modelled cumulative (time-integrated) glacial erosion
          potential along a Rhine Glacier transect (upper panels dashed line).}
      \label{fig:powerlaws}
    \end{figure*}

>>>>

    The choice of erosion law significantly impacts the results
    (Fig.~\ref{fig:powerlaws}). Our default erosion law, based on quantified
    sediment yields from Patagonian and Antarctic Peninsula tidewater glaciers
    \citep[${\dot{e}=5.2\times10^{-8}u_\mathrm{b}^{2.34}}$,][]{Koppes.etal.2015}
    yields a moderate and strongly localized cumulative erosion potential
    (Figs~\ref{fig:cumulative} and~\ref{fig:powerlaws}a). The erosion law
    based on measured suspended sediment load from Franz-Joseph Glacier
    \citep[${\dot{e}=2.7\times10^{-7}u_\mathrm{b}^{2.02}}$,][]{Herman.etal.2015}
    also yields a strongly localized, and yet much higher, modelled erosion
    potential (Fig.~\ref{fig:powerlaws}b). The erosion law based on a global
    compilation of glacier velocity and erosion rates
    \citep[${\dot{e}=1.665\times10^{-1}u_\mathrm{b}^{0.6459}}$,][]{Cook.etal.2020}
    results in a similarly high values, but a flatter (less localized) pattern
    of cumulative erosion potential (Fig.~\ref{fig:powerlaws}c).

    It may seem surprising that the erosion law calibrated on Patagonian and
    Antarctic Peninsula tidewater glaciers \citep{Koppes.etal.2015} yields a
    more realistic cumulative erosion potential than those based on terrestrial
    glaciers \citep{Herman.etal.2015, Cook.etal.2020}, themselves resulting
    in cumulative erosion potential values in the order of kilometres and thus
    higher than the Alpine relief itself. However, during the Last Glacial
    Maximum and much of the last glacial cycle, Alpine paleoglaciers were
    closer in size and climatic context to the present-day glaciers of
    Patagonia and the Antarctic Peninsula \citep{Koppes.etal.2015} than to
    Franz-Joseph Glacier \citep{Herman.etal.2015} and many of the glaciers
    included in the global compilation by \citet{Cook.etal.2020}. Atmospheric
    temperatures have been found to be a strong driver of erosion in studies
    involving several glaciers \citep{Koppes.etal.2015, Cook.etal.2020}.

    That said, the magnitude of the modelled erosion rates can hardly be used
    as a criteria to rule out this or that erosion law for the glacial Alps, as
    all these laws are affected with very large uncertainties. It can not be
    excluded that a linear or sub-linear erosion-law
    \citep[e.g.][]{Cook.etal.2020} would lead to a less localized pattern of
    erosion (Fig.~\ref{fig:powerlaws}c), and a stronger correlation between
    modelled ice volume and erosion \note{not shown, should it be?}. Even
    then, the intrinsically large variations in glacier sliding velocities
    yields modelled glacial-cycle erosion potential two to three orders of
    magnitude higher in the Alpine valley than near the mountain tops.

@@

    \begin{figure*}
      \centerline{\includegraphics{alpero_powerlaws}}
      \caption{%
        Modelled cumulative (time-integrated) glacial erosion potential for
        three different erosion laws published by
        \textbf{(a)} \citet{Koppes.etal.2015},
          ${5.2 \times 10^{-8} u_\mathrm{b}^{2.34}}$ (same as
          Fig.~\ref{fig:cumulative}a but with a different colour scale),
        \textbf{(b)} \citet{Herman.etal.2015},
          ${2.7 \times 10^{-7} u_\mathrm{b}^{2.02}}$, and
        \textbf{(c)} \citet{Cook.etal.2020},
          ${1.665 \times 10^{-1} u_\mathrm{b}^{0.6459}}$.
        \textbf{(d)} Corresponding erosion power laws, and
        \textbf{(e)} modelled cumulative (time-integrated) glacial erosion
          potential along a Rhine Glacier transect (upper panels dashed line).}
      \label{fig:powerlaws}
    \end{figure*}
